---
title: "BigDataFinal"
author: "brian rochford"
date: "5/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, warning=FALSE}
library(haven)
library(usmap)
library(tidyverse)
library(stargazer)
library(kableExtra)
library(AER)
library(wooldridge)
library(plm)
library(stargazer)
library(haven)
library(labelled)
library(lubridate)
library(XLConnect)
library(rJava)
library(xlsx)
library(readxl)
library(ggplot2)
library(gtsummary)
```
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(wooldridge)
library(stargazer)
library(kableExtra)
library(haven)
library(ggplot2)
library(dplyr)
library(purrr)
library(reshape2)
library(growthcurver)
library(lubridate)
library(XLConnect)
library(lme4)
library(car)
library(rJava)
library(xlsx)
library(readxl)
library(ggplot2)
library(zoo)
library(lubridate)
library(hydroTSM)
```
#```{r }

#df <- read_csv("/Users/brianrochford/Downloads/fullyNormAndCleanBigData.csv")
##df
##```df <- read_csv("/Users/brianrochford/Downloads/BDOnlyNyc.csv")
#first model ran


```{r }

df <- read_csv("/Users/brianrochford/Downloads/fullycleanedBigData.csv")
df
```
##second model with 25-75 percentile and NA dups dropped
#```{r }

#df <- read_csv("/Users/brianrochford/Downloads/fullycleanedBigData.csv")
#df
#```

```{r}
year_set = df %>% distinct(year)
df = df %>% mutate(
year2017 = 0,
year2018 = 0,
year2019 = 0,
year2020 = 0,
year2021 = 0,
)
y2017 = df %>% filter(year == 2017) %>% mutate(year2017 = 1)
y2018 = df %>% filter(year == 2018) %>% mutate(year2018 = 1)
y2019 = df %>% filter(year == 2019) %>% mutate(year2019 = 1)
y2020 = df %>% filter(year == 2020) %>% mutate(year2020 = 1)
y2021 = df %>% filter(year == 2021) %>% mutate(year2021 = 1)
new_data = y2017 %>%
rbind(y2018) %>%
rbind(y2019) %>%
rbind(y2020) %>%
rbind(y2021)
```
```{r}
borough_set = new_data %>% distinct(borough)
new_data = new_data %>% mutate(
boroughManhattan = 0,
boroughQueens = 0,
boroughBrooklyn = 0,
boroughBronx = 0,
)
by2017 = new_data %>% filter(borough =="MANHATTAN") %>% mutate(boroughManhattan = 1)
by2018 = new_data %>% filter(borough == "QUEENS") %>% mutate(boroughQueens = 1)
by2019 = new_data %>% filter(borough == "BROOKLYN") %>% mutate(boroughBrooklyn = 1)
by2020 = new_data %>% filter(borough == "BRONX") %>% mutate(boroughBronx = 1)
new_data = by2017 %>%
rbind(by2018) %>%
rbind(by2019) %>%
rbind(by2020) 
```



```{r }
model1 <- lm(TotalHourlyCompensatio ~ YearsEmployedAtAgency + year2018 + year2019 + year2020 + year2021 + boroughManhattan + boroughBrooklyn + boroughBronx,
data = new_data, na.action = na.omit)
summary(model1)

```
```{r }
plot(model1, which = 4)

```
```{r}
df_ts2 <- ts(df, start = c(2017, 1), frequency = 12)
df_ts2
```
```{r message=FALSE, warning=FALSE}
serial_corrTest = serialCorrelationTest(model1, test = "rank.von.Neumann", alternative = "two.sided", conf.level = 0.95)
summary(serial_corrTest)
```




```{r results= 'asis'}
stargazer(model1, style = "aer", type = "html", title = "Wage Regression by Year and Borough", covariate.labels = c("YearsEmployed", "year2018", "year2019", "year2020","year2021", "boroughManhattan","boroughBrooklyn", "boroughBronx"), dep.var.labels = c("HourlyWage"), column.labels = c("Region Year Dummies compared to Queens and 2017"))
```

```{r }
model_copy <- lm(TotalHourlyCompensatio ~ YearsEmployedAtAgency + year2018 + year2019 + year2020 + year2021 + boroughManhattan + boroughBrooklyn + boroughBronx,
data = new_data, na.action = na.omit)
summary(model_copy)

```

```{r}
model5 = lm(TotalHourlyCompensatio ~ 0 + YearsEmployedAtAgency + I(boroughManhattan * year2018) + I(boroughQueens * year2018) + I(boroughBrooklyn * year2018),
data = new_data, na.action = na.omit)
summary(model5)
```

```{r}
model2019 = lm(TotalHourlyCompensatio ~ 0 + YearsEmployedAtAgency + I(boroughManhattan * year2019) + I(boroughQueens * year2019) + I(boroughBrooklyn * year2019),
data = new_data, na.action = na.omit)
summary(model2019)
```
```{r}
model2020 = lm(TotalHourlyCompensatio ~ 0 + YearsEmployedAtAgency + I(boroughManhattan * year2020) + I(boroughQueens * year2020) + I(boroughBrooklyn * year2020),
data = new_data, na.action = na.omit)
summary(model2020)
```

```{r}
model2021 = lm(TotalHourlyCompensatio ~ 0 + YearsEmployedAtAgency + I(boroughManhattan * year2021) + I(boroughQueens * year2021) + I(boroughBrooklyn * year2021),
data = new_data, na.action = na.omit)
summary(model2021)
```



```{r results= 'asis'}
stargazer(model5, model2019, model2020, model2021, style = "aer", type = "html", title = "Wage Regression by Year and Borough", covariate.labels = c("YearsEmployed", "I(boroughManhattan * year2018)", "I(boroughQueens * year2018)","I(boroughBrooklyn * year2018)","I(boroughManhattan * year2019)", "I(boroughQueens * year2019)","I(boroughBrooklyn * year2019)","I(boroughManhattan * year2020)", "I(boroughQueens * year2020)","I(boroughBrooklyn * year2020)","I(boroughManhattan * year2021)", "I(boroughQueens * year2021)","I(boroughBrooklyn * year2021)"), dep.var.labels = c("HourlyWage"), column.labels = c("2018", "2019", "2020", "2021"))
```







